Splitting of the homology of the punctured mapping class group

### 2020-05-01 arxiv version/HBraidSurf.tex
Final version ready to appear in JTop

% isomorphism of bigraded vector spaces
% \begin{equation}
% \label{eq:genuszero}
% \bigoplus_{m\geq 0} H_*\pa{C_m(\D)}\simeq \Z_2[Q^j\epsilon\,|\, j\geq 0].
% \end{equation}
% This isomorphism was essentially proved by Fuchs in \ref{Fuchs:CohomBraidModtwo}. Moreover the left hand
% side is the homology of the space $\coprod_{m\geq 0} C_m(\D)$, which has a natural structure
% of algebra over the operad $E^2$ of little $2$-cubes. The isomorphism in equation \eqref{eq:genuszero}
% is an isomorphism of rings (using the Pontryagin product on the left hand side), and $Q$ represents here
% the Araki-Kudo-Dyer-Lashof operation $Q\colon H_*(X;\Z_2)\to H_{2*+1;\Z_2}(X)$,
% which is defined for any $E^2$-algebra $X$; $\epsilon\in H_0(C_1(\D))$ is the generator.
% This explains our notation; we refer to Cohen \cite[Chap.3]{CLM} for more details.

In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2$-vector spaces}
as in Theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
$\gg$.
%In these two sections we abbreviate $\S=\sg$.

Since we work with coefficients in a field, it is equivalent to compute homology or cohomology,
and in this section we will prefer to compute $H^*(\cms)$ for all bigradings $(*,m)$.
% the dual graded vector space $H_*(\cms)$ is abstractly isomorphic,
% i.e. it has the same dimension.

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defTS.pdf}
 \caption{The space $Q(\S)$ for $\S=\Sigma_{3,1}$. The black pentagons represent the point $\cP_{even}$, whereas the black
 squares represent $\cP_{odd}$.}
\label{fig:defTS}
\end{figure}

Our next aim is to define a structure of $CW$-complex on the space $\cms^{\infty}$:
the only $0$-cell will be the point $\infty$, whereas the other cells will be introduced
in the following definition.
\begin{defn}
\label{defn:ehopen}
A \emph{tuple} $\tup$ is a choice of the following set of data:
 \begin{itemize}
  \item a natural number $0\leq l\leq m$;
  \item a vector $\ux=(x_1,\dots, x_l)$ of integers $\geq 1$;
  \item vectors $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ of integers $\geq 0$,
 \end{itemize}
satisfying the following equality
\[
 m=\sum_{i=1}^lx_i+\sum_{i=1}^g(u_i+v_i).
\]
% We will generically use the letter $m$ to denote any of the numbers $l$, $x_i$,$y_i$,$z$,$u_i$,$v_i$ or $w_i$
% in the above sum, so $m\in\Z$; instead .
% Similarly we will use the letter $\M$ to denote any of the intervals $\U_i,\V_i,\W_i$;
% they will correspond to the letters $U_i,V_i,W_i$, so we also define a subset
% $\bsymb=\set{(U_i,V_i)_{i\leq g},(W_i)_{i\leq n-1}}\subset\symb$.
In symbols we write $\tup=(l,\ux,\uu,\uv)$. We omit from the notation $\ux$ if it vanishes: this happens precisely when $l=0$.
The \emph{dimension} of $\tup$ is defined as $m+l$.

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{figures/defetup.pdf}
 \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,2,0),(0,3,2))$. Points
 lying on the $\U_i$'s and the $\V_i$'s are depicted twice.}
\label{fig:defetup}
\end{figure}

@@

% We want to show that the collection of the $e_h$ for varying $h$, together with
% the 0-cell $*$, give a cell decomposition of $C_k(\S)^{\infty}$. Denote by $\Delta^h$ the following product of simplices
\[
 \Delta^{\tup}\colon =
 \Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}},
%  =\prod_{m\in\symb_l}\Delta^m.
\]
where the simplex $\Delta^r$ is the subspace of $[0,1]^r$ of sequences $0\leq \tau_1\leq\dots\leq\tau_r\leq 1$
(the numbers $\tau_1,\dots,\tau_r$ are the \emph{local coordinates} of the simplex). The homeomorphism is given
as follows:
\begin{itemize}
 \item the local coordinates of the $\Delta^l$-factor correspond to the positions $s_1,\dots,s_l$ of the vertical
 lines in $(0,1)^2$ containing points of the configuration;
 \item the local coordinates of the $\Delta^{x_i}$-factor correspond to the positions of the $x_i$ points
 lying on the vertical line $\set{s_i}\times(0,1)$;
 \item the local coordinates of the $\Delta^{u_i}$-factor correspond to the positions of the $u_i$ points
 lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}$-factor,
 with $v_i$ and $\V_i$ replacing $u_i$ and $\U_i$.
\end{itemize}
Note that the dimension of $\Delta^{\tup}$ agrees with the dimension of $\tup$ from Definition \ref{defn:ehopen}.
The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^{\infty}$ extends to a continuous map
$\Phi^{\tup}\colon\Delta^{\tup}\to\cms^{\infty}$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
all subspaces $e^{\tup'}$ for tuples $\tup'$ of \emph{lower} dimension than $\tup$, together with the $0$-cell $\infty$.

@@

We compute the \emph{reduced} cellular chain complex of $\cms^{\infty}$
with coefficients in $\Z_2$. It is a chain complex in $\Z_2$-vector spaces with basis given
by tuples $\tup$, which correspond to the cells $e^{\tup}$.
\begin{lem}
\label{lem:doperatoropenmodtwo}
Let $\tup=(l,\ux,\uu, \uv)$ and
$\tup'=(l-1,\ux',\uu',\uv')$
be tuples in consecutive dimensions $m+l$ and $m+l-1$. Denote by $[\tup\colon\tup']\in\Z_2$ the coefficient
of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$.
Then $[\tup:\tup']=0$ unless one, and exactly one, of the following situations occurs:
\begin{itemize}
 \item $l\geq 2$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, setting $x'_i=x_i+x_{i+1}$
 for one value $1\leq i\leq l-1$, and shifting the values
 $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case we say that $\tup'$ is an \emph{inner boundary} of $\tup$, and
 we have
 \[
  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2.
 \]
%  This binomial coefficient may be anyway equal to $0\in\Z_2$.
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_1=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and shifting the values
 $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case we say that $\tup'$ is a \emph{left, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_l=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
 In this case we say that $\tup'$ is a \emph{right, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
%  This product of binomial coefficients may be anyway equal to $0\in\Z_2$.
\end{itemize}

@@

 The only possibility in which the same cell $e^{\tup'}$ is hit by different faces $\partial_i^{\hor}\Delta^{\tup}$
 and $\partial_j^{\hor}\Delta^{\tup}$ is the one described in the remark preceding this proof: in this case
 there are two equal contributions $[\partial_i^{\hor}\tup\colon\tup']$ and $[\partial_j^{\hor}\tup\colon\tup']$
 that cancel each other.
%  We are working in $\Z_2$ so we do not have to be careful about orientations of cells
%  and therefore about signs.
\end{proof}

@@

The associated \emph{symmetric chain} in $\tCh_*(C_p(\D)^{\infty})$, denoted by $\kappa(\ualpha)$,
is the sum of all tuples $\tup=\pa{l,(x_i)_{i\leq l}}$ such that
\begin{itemize}
 \item $l=\sum_{j\geq 1}\alpha_j$;
 \item every $x_i$ is a power of 2;
 \item for all $j\geq 0$ there are exactly $\alpha_j$ indices $i$ such that $x_i=2^j$.
\end{itemize}
% Here and in the following it is understood that $\alpha_j=0$ for all but finitely many indices $j$.
A symmetric chain $\kappa(\ualpha)$ is a cycle in the chain complex $\tCh_*(C_p(\D)^{\infty})$ (see \cite{Fuchs:CohomBraidModtwo}).
We denote by $[\kappa(\ualpha)]\in\tH_{p+l}(C_p(\D)^{\infty})$ the associated homology class.
\end{defn}
In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\D)^{\infty})$
is given by the collection of all classes $[\kappa(\ualpha)]$ associated to sequences $\ualpha=(\alpha_j)_{j\geq 0}$
which satisfy the equality $p=\sum_{j\geq 0}\alpha_j2^j$.

@@

% This notation is compatible, after interpreting equation \eqref{eq:genuszero} as an isomorphism
% of rings, with the results of \cite[Chap.3]{CLM}: the monomials constructed there, using
% the Pontryagin product and the operation $Q$, are really the dual basis of the basis of
% symmetric chains for cohomology.

We will not need this finer result in this article, so in the following the expression
$Q^{\ualpha}$ will only denote
the (unique) homology class in $H_{p-l}(C_p(\D)$ such that the following holds:
for all $\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\sum_{j\geq 0}\alpha'_j=p$, the
algebraic intersection between $Q^{\ualpha}\epsilon\in H_{p-l}(C_p(\D))$
and $[\kappa(\ualpha')]\in\tH_{p+l}(C_p(\D)^{\infty})$
is $1\in\Z_2$ if and only $\ualpha=\ualpha'$.
% i.e. we \emph{define} monomials as the dual basis of the basis of symmetric chains, and keep considering
% \eqref{eq:Cohen} only as an isomorphism of bigraded vector spaces.

@@

The spectral sequence considered above collapses on its first page and
% $H_*(\cms^{\infty})$
% has a $\Z_2$-basis given by generalised symmetric chains.
we have the following lemma:
\begin{lem}
\label{lem:gensymchain}
The homology $H_*(\cms^{\infty})$ has a graded basis given by the classes $[\kappa(p,\ualpha,\uu,\uv)]$
associated to generalised symmetric chains of weight $m$.
% this basis is bijection with
% the choices of the following set of data:
% \begin{itemize}
%  \item a number $0\leq p\leq m$;
%  \item a splitting $m-p=\sum_{i=1}^g (u_i+v_i)$ with $u_i,v_i\geq 0$;
%  \item a splitting $p=\sum_{j=0}^{\infty}\alpha_j2^j$ with $\alpha_j\geq 0$.
% \end{itemize}
% The homology class $[p,(u_i,v_i),(\alpha_j)]\in H_*(\cms^{\infty})$ has homological degree
% $m+\sum_{j=1}^{\infty}\alpha_j$.
\end{lem}
% To interpret this class
% as a monomial in the tensor product of Theorem \ref{thm:Hbms*as*ggrep} we need the following definition.

% We are almost ready to prove an isomorphism of bigraded $\Z_2$-vector spaces as in Theorem \ref{thm:Hbms*as*ggrep}.

\begin{defn}
 \label{defn:dualHbasis}
We can see the $\U_i$'s and $\V_i$'s as properly embedded $1$-manifolds in $\mrS$;
by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^{\infty}}\simeq H^1(\S)$,
and in particular they form a basis
of the latter cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.

% We fix simple closed curves $\u_i,\v_i$ on $\S$, for $1\leq i\leq g$, whose fundamental classes
% $[\u_i],[\v_i]$ form the dual basis of $H_1(\S)$. We assume that, apart from the following necessary exceptions,
% all curves $\U_i,\V_i,\u_i,\v_i$ for $1\leq i\leq g$ are disjoint:
% \begin{itemize}
%  \item $\u_i$ and $\v_i$ intersect once, transversely;
%  \item $\u_i$ and $\U_i$ intersect once, transversely;
%  \item $\v_i$ and $\V_i$ intersect once, transversely;
%  \end{itemize}
% SEE PICTURE.
\end{defn}
We establish a bijection between monomials in the tensor product of Theorem \ref{thm:Hbms*as*ggrep}
and the basis of $H^*(\cms)\simeq H_*(\cms^{\infty})$ in Lemma \ref{lem:gensymchain}:
the class 
\[
[\kappa(p,\ualpha,\uu,\uv)]\in H_{m+\sum\alpha_j}(\cms^{\infty})\simeq H^{m-\sum\alpha_j}(\cms)
\]
is associated with the monomial
\[
Q^{\ualpha}\epsilon\cdot\u^{\uu}\cdot\v^{\uv}\colon =\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g(\u_i^{u_i}\v_i^{v_i}),
\]
where $\ualpha=(\alpha_j)_{j\geq 0}$, $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$.

>>>>

In this section we will prove that there is an isomorphism of \emph{bigraded $\Z_2$-vector spaces}
as in Theorem \ref{thm:Hbms*as*ggrep}; in the next section we will deal with the action of
$\gg$.

Since we work with coefficients in a field, it is equivalent to compute homology or cohomology,
and in this section we will prefer to compute $H^*(\cms)$ for all bigradings $(*,m)$.

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{defTS.pdf}
 \caption{The space $Q(\S)$ for $\S=\Sigma_{3,1}$. The black pentagons represent the point $\cP_{even}$, whereas the black
 squares represent $\cP_{odd}$.}
\label{fig:defTS}
\end{figure}

Our next aim is to define a structure of $CW$-complex on the space $\cms^{\infty}$:
the only $0$-cell will be the point $\infty$, whereas the other cells will be introduced
in the following definition.
\begin{defn}
\label{defn:ehopen}
A \emph{tuple} $\tup$ is a choice of the following set of data:
 \begin{itemize}
  \item a natural number $0\leq l\leq m$;
  \item a vector $\ux=(x_1,\dots, x_l)$ of integers $\geq 1$;
  \item vectors $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$ of integers $\geq 0$,
 \end{itemize}
satisfying the following equality
\[
 m=\sum_{i=1}^lx_i+\sum_{i=1}^g(u_i+v_i).
\]
In symbols we write $\tup=(l,\ux,\uu,\uv)$. We omit from the notation $\ux$ if it vanishes: this happens precisely when $l=0$.
The \emph{dimension} of $\tup$ is defined as $m+l$.

@@

\begin{figure}\centering
 \includegraphics[scale=0.7]{defetup.pdf}
 \caption{A configuration in the subspace $e^{\tup}\subset C_{14}(\S)$, for the tuple $\tup=(3,(3,1,2),(1,2,0),(0,3,2))$. Points
 lying on the $\U_i$'s and the $\V_i$'s are depicted twice.}
\label{fig:defetup}
\end{figure}

@@

\[
 \Delta^{\tup}\colon =
 \Delta^l\times\prod_{i=1}^l\Delta^{x_i}\times\prod_{i=1}^g\pa{\Delta^{u_i}\times\Delta^{v_i}},
\]
where the simplex $\Delta^r$ is the subspace of $[0,1]^r$ of sequences $0\leq \tau_1\leq\dots\leq\tau_r\leq 1$
(the numbers $\tau_1,\dots,\tau_r$ are the \emph{local coordinates} of the simplex). The homeomorphism is given
as follows:
\begin{itemize}
 \item the local coordinates of the $\Delta^l$-factor correspond to the positions $s_1,\dots,s_l$ of the vertical
 lines in $(0,1)^2$ containing points of the configuration;
 \item the local coordinates of the $\Delta^{x_i}$-factor correspond to the positions of the $x_i$ points
 lying on the vertical line $\set{s_i}\times(0,1)$;
 \item the local coordinates of the $\Delta^{u_i}$-factor correspond to the positions of the $u_i$ points
 lying on $\U_i$, which is canonically identified with $(0,1)$; similarly for the $\Delta^{v_i}$-factor,
 with $v_i$ and $\V_i$ replacing $u_i$ and $\U_i$.
\end{itemize}
Note that the dimension of $\Delta^{\tup}$ agrees with the dimension of $\tup$ from Definition \ref{defn:ehopen}.
The embedding $\mathring{\Delta}^{\tup}\cong e^{\tup}\hookrightarrow \cms^{\infty}$ extends to a continuous map
$\Phi^{\tup}\colon\Delta^{\tup}\to\cms^{\infty}$, so that the image of $\partial\Delta^{\tup}$ is contained in the union of
all subspaces $e^{\tup'}$ for tuples $\tup'$ of \emph{lower} dimension than $\tup$, together with the $0$-cell $\infty$.

@@

We compute the \emph{reduced} cellular chain complex of $\cms^{\infty}$
with coefficients in $\Z_2$. It is a chain complex in $\Z_2$-vector spaces with basis given
by tuples $\tup$, which correspond to the cells $e^{\tup}$.
\begin{lem}
\label{lem:doperatoropenmodtwo}
Let $\tup=(l,\ux,\uu, \uv)$ and
$\tup'=(l-1,\ux',\uu',\uv')$
be tuples in consecutive dimensions $m+l$ and $m+l-1$. Denote by $[\tup\colon\tup']\in\Z_2$ the coefficient
of $\tup'$ in $\partial \tup$ in the reduced chain complex $\tilde\Ch_*(\cms^{\infty})$.
Then $[\tup:\tup']=0$ unless one, and exactly one, of the following situations occurs:
\begin{itemize}
 \item $l\geq 2$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, setting $x'_i=x_i+x_{i+1}$
 for one value $1\leq i\leq l-1$, and shifting the values
 $x'_j=x_{j+1}$ for $i+1\leq j\leq l-1$. In this case we say that $\tup'$ is an \emph{inner boundary} of $\tup$, and
 we have
 \[
  [\partial \tup\colon\tup']=\binom{x_i+x_{i+1}}{x_i}\in\Z_2,
 \]
 where $\binom{x_i+x_{i+1}}{x_i}$ denotes the binomial coefficient.
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_1$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_1=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and shifting the values
 $x'_j=x_{j+1}$ for $1\leq j\leq l-1$. In this case we say that $\tup'$ is a \emph{left, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
 \item $l\geq 1$ and $\tup'$ is obtained from $\tup$ by decreasing $l$ by 1, choosing a splitting of $x_l$
 in integers $\tildeu_i,\tildev_i\geq 0$
 \[
  x_l=\sum_{i=1}^g \pa{\tildeu_i+\tildev_i},
 \]
 setting $u'_i=u_i+\tildeu_i$ and $v'_i=v_i+\tildev_i$ for all $1\leq i\leq g$ and keeping $x'_i=x_i$ for all $1\leq i\leq l-1$.
 In this case we say that $\tup'$ is a \emph{right, outer boundary} of $\tup$, and we have
 \[
  [\partial \tup\colon\tup']=\prod_{i=1}^g\binom{u_i+\tildeu_i}{u_i}\binom{v_i+\tildev_i}{v_i}\in\Z_2.
 \]
\end{itemize}

@@

 The only possibility in which the same cell $e^{\tup'}$ is hit by different faces $\partial_i^{\hor}\Delta^{\tup}$
 and $\partial_j^{\hor}\Delta^{\tup}$ is the one described in the remark preceding this proof: in this case
 there are two equal contributions $[\partial_i^{\hor}\tup\colon\tup']$ and $[\partial_j^{\hor}\tup\colon\tup']$
 that cancel each other.
\end{proof}

@@

The associated \emph{symmetric chain} in $\tCh_*(C_p(\D)^{\infty})$, denoted by $\kappa(\ualpha)$,
is the sum of all tuples $\tup=\pa{l,(x_i)_{i\leq l}}$ such that
\begin{itemize}
 \item $l=\sum_{j\geq 1}\alpha_j$;
 \item every $x_i$ is a power of 2;
 \item for all $j\geq 0$ there are exactly $\alpha_j$ indices $i$ such that $x_i=2^j$.
\end{itemize}
A symmetric chain $\kappa(\ualpha)$ is a cycle in the chain complex $\tCh_*(C_p(\D)^{\infty})$ (see \cite{Fuchs:CohomBraidModtwo}).
We denote by $[\kappa(\ualpha)]\in\tH_{p+l}(C_p(\D)^{\infty})$ the associated homology class.
\end{defn}
In \cite{Fuchs:CohomBraidModtwo} Fuchs shows that a graded basis for $\tH_*(C_p(\D)^{\infty})$
is given by the collection of all classes $[\kappa(\ualpha)]$ associated to sequences $\ualpha=(\alpha_j)_{j\geq 0}$
which satisfy the equality $p=\sum_{j\geq 0}\alpha_j2^j$.

@@

We will not need this finer result in this article, so in the following the expression
$Q^{\ualpha}$ will only denote
the (unique) homology class in $H_{p-l}(C_p(\D)$ such that the following holds:
for all $\ualpha'=(\alpha'_j)_{j\geq 0}$ with $\sum_{j\geq 0}\alpha'_j=p$, the
algebraic intersection between $Q^{\ualpha}\epsilon\in H_{p-l}(C_p(\D))$
and $[\kappa(\ualpha')]\in\tH_{p+l}(C_p(\D)^{\infty})$
is $1\in\Z_2$ if and only $\ualpha=\ualpha'$.

@@

The spectral sequence considered above collapses on its first page and
we have the following lemma:
\begin{lem}
\label{lem:gensymchain}
The homology $H_*(\cms^{\infty})$ has a graded basis given by the classes $[\kappa(p,\ualpha,\uu,\uv)]$
associated to generalised symmetric chains of weight $m$.
\end{lem}

\begin{defn}
 \label{defn:dualHbasis}
We can see the $\U_i$'s and $\V_i$'s as properly embedded $1$-manifolds in $\mrS$;
by Poincaré-Lefschetz duality they represent classes in $\tH_1\pa{(\mrS)^{\infty}}\simeq H^1(\S)$,
and in particular they form a basis
of the latter cohomology group. We call $\u_i,\v_i\in H_1\pa{\mrS}$ the dual basis.
\end{defn}
We establish a bijection between monomials in the tensor product of Theorem \ref{thm:Hbms*as*ggrep}
and the basis of $H^*(\cms)\simeq H_*(\cms^{\infty})$ in Lemma \ref{lem:gensymchain}:
the class 
\[
[\kappa(p,\ualpha,\uu,\uv)]\in H_{m+\sum\alpha_j}(\cms^{\infty})\simeq H^{m-\sum\alpha_j}(\cms)
\]
is associated with the monomial
\[
Q^{\ualpha}\epsilon\cdot\u^{\uu}\cdot\v^{\uv}\colon =\prod_{j=1}^{\infty}(Q^j\epsilon)^{\alpha_j}\otimes \prod_{i=1}^g(\u_i^{u_i}\v_i^{v_i}),
\]
where $\ualpha=(\alpha_j)_{j\geq 0}$, $\uu=(u_1,\dots,u_g)$ and $\uv=(v_1,\dots,v_g)$.

### 2020-05-01 arxiv version/Introduction.tex
Final version ready to appear in JTop

Forgetting the punctures gives a surjective map $\ggm\to\gg$ with kernel $\bms$,
the $m$-th \emph{braid group} of the surface $\sg$. We obtain the Birman exact sequence (see \cite{Birman:mcgbr})
\begin{equation}
\label{eq:Birman}
1\to\bms\to\ggm\to\gg\to 1.
\end{equation}
% This short exact sequence corresponds to 
% \begin{equation}
%  \cms\to\mgm\to\mg.
% \end{equation}
% Here $\cms$ is the \emph{unordered} configuration space of $m$ points in the interior of $\Sigma$, i.e.
% \begin{equation}
%  \cms=\fms/\S_m=\set{(p_1,\dots,p_m)\in\pa{\mathring\sg}^m|p_i\neq p_j \;\forall i\neq j}/\S_m,
% \end{equation}
% the quotient of the \emph{ordered} configuration space of $m$ points in the interior of $\Sigma$ by the natural
% action of the symmetric group $\S_m$ that permutes the (ordered list of) points in the configuration; $\mg$ and $\mgm$
% are the moduli spaces of Riemann structures on $\sg$ and $\sgm$ respectively.

@@

The main result of this article is that this spectral sequence collapses in its second page.
\begin{thm}
\label{thm:main}
For all $l\geq 0$ there is an isomorphism of vector spaces
\begin{equation}
\label{eq:main}
H_l\pa{\ggm;\Z_2}\cong \bigoplus_{k+q=l} H_k\pa{\gg;H_q\pa{\bms;\Z_2}}.
\end{equation}
\end{thm}
Thus the computation of $H_*\pa{\ggm;\Z_2}$ reduces to the computation of the homology of $\gg$ with
twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\gg$-representation
splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action: this is done in
Theorem \ref{thm:Hbms*as*ggrep}, which together with Theorem \ref{thm:main} is the main result of the article.
% in section \ref{sec:HBraidSurf}).

%In particular this representation is symplectic: this had already benn stated in \cite{LM}.

% The group $\gg$ is significantly smaller than $\ggm$, so the
% computation of the right-hand side of equation \eqref{eq:main}
% may be more tractable.
%WE DO THIS COMPUTATION FOR $g=1$.

The strategy of the proof does not generalize to fields of characteristic different from 2 or to the \emph{pure}
mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In section \ref{sec:rational}
we describe in detail a counterexample with coefficients in $\Q$, which can be generalized both to coefficients
in a field $\F_p$ of odd characteristic and to the pure mapping class group.

>>>>

Forgetting the punctures gives a surjective map $\ggm\to\gg$ with kernel $\bms$,
the $m$-th \emph{braid group} of the surface $\sg$. We obtain the Birman exact sequence (see \cite{Birman:mcgbr})
\begin{equation}
\label{eq:Birman}
1\to\bms\to\ggm\to\gg\to 1.
\end{equation}

@@

The main result of this article is that this spectral sequence collapses in its second page.
\begin{thm}
\label{thm:main}
For all $l\geq 0$ there is an isomorphism of vector spaces
\begin{equation}
\label{eq:main}
H_l\pa{\ggm;\Z_2}\cong \bigoplus_{k+q=l} H_k\pa{\gg;H_q\pa{\bms;\Z_2}}.
\end{equation}
\end{thm}
Thus the computation of $H_*\pa{\ggm;\Z_2}$ reduces to the computation of the homology of $\gg$ with
twisted coefficients in the representation $H_*\pa{\bms;\Z_2}$. We will see that this $\gg$-representation
splits as a direct sum of symmetric powers of $H_1(\sg)$ with the symplectic action: this is done in
Theorem \ref{thm:Hbms*as*ggrep}, which together with Theorem \ref{thm:main} is the main result of the article.

The strategy of the proof does not generalize to fields of characteristic different from 2 or to the \emph{pure}
mapping class group, in which we consider only diffeomorphisms of $\sgm$ that fix all punctures. In Section \ref{sec:rational}
we describe in detail a counterexample with coefficients in $\Q$, which can be generalized both to coefficients
in a field $\F_p$ of odd characteristic and to the pure mapping class group.
